---
title: 'Exploratory Cost-Effectiveness Analysis of Treatment Based on New Diagnostic Tests vs Standard of Care in Microsatellite Stable RAS Mutant Metastatic Colorectal Cancer Patients'
author: "Jonathan Briody (1) | Kathleen Bennett (1)"
output:
  html_document: default
  pdf_document: default
---

(1) Data Science Centre, RCSI University of Medicine and Health Sciences, Dublin, Ireland


**Correspondence:**

Jonathan Briody, PhD, RCSI University of Medicine and Health Sciences, Beaux Lane House, Lower Mercer St, Dublin 2, Ireland. Email: jonathanbriody@rcsi.ie


**Funding Information:**

This project has received funding from the European Union’s Horizon 2020 research and innovation programme under grant agreement No 754923. The material presented and views expressed here are the responsibility of the author(s) only. The EU Commission takes no responsibility for any use made of the information set out.

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval = T)
# To knit this document every time it is run, you can change `eval` to `TRUE` in the above.
```


```{r}
rm(list = ls())  
# clear memory (removes all the variables from the work space)
```

# 01 Load packages

```{r}
if (!require('pacman')) install.packages('pacman'); library(pacman) 
# use this package to conveniently install other packages
# load (install if required) packages from CRAN
p_load("diagram", "dampack") 
# install_github("DARTH-git/darthtools", force = TRUE) Uncomment if there is a newer version
p_load_gh("DARTH-git/darthtools")
```

# 02 Load functions

```{r}
# all functions are in the darthtools package
```

# 03 Input model parameters

```{r}
## General setup
# Here we define all our model parameters, so that we can call on these parameters later during our model:

n_cycles        <- 60                            
# We set the number of cycles to 60 to reflect 5 years broken down into monthly cycles
v_names_cycles  <- paste("cycle", 0:n_cycles)    
# So here, we just name each cycle by the cycle its on, going from o up to the number of cycles there are, here 60.
v_names_states  <- c("Healthy", "Sick", "Dead")  
# These are the health states in our model, PFS, OS, Death.
n_states        <- length(v_names_states)        
# We're just taking the number of health states from the number of names we came up with, i.e. 3 names to reflect three health states 

# Discounting factors
d_c             <- 0.03                          
# discount rate for costs
d_e             <- 0.03                          
# discount rate for QALYs

# Strategy names
v_names_str     <- c("Standard of Care",         
                     "Treatment A", 
                     "Treatment B")  
# store the strategy names
n_str           <- length(v_names_str)           
# number of strategies


## Transition probabilities
p_HS_SoC  <- 0.05  # probability of becoming sick when healthy, conditional on surviving, under standard of care
p_HS_trtA <- 0.04  # probability of becoming sick when healthy, conditional on surviving, under treatment A
p_HS_trtB <- 0.02  # probability of becoming sick when healthy, conditional on surviving, under treatment B
p_SD      <- 0.1   # probability of dying          
p_HD      <- 0.01  # probability of dying when healthy



# To add age specific mortality to our model, we would use this #03 input model parameters of:

# "C:\Users\Jonathan\OneDrive - Royal College of Surgeons in Ireland\COLOSSUS\Training Resources\Decision Modeling for Public Health_DARTH\5_Nov_29\4_Cohort state-transition models (cSTM) - time-dependent models_material\Markov_3state_time"

# with the 55 minute mark of: C:\Users\Jonathan\OneDrive - Royal College of Surgeons in Ireland\COLOSSUS\Training Resources\Cost-Effectiveness and Decision Modeling using R Workshop _ DARTH\August_24\Live Session Recording\Live Session Recording August 24th.mp4

# and this would allow us to create a vector of transition probabilities for p_HD above, i.e., from healthy to dead, that is a little bit larger at each cycle, starting at our chosen minimum value at the first cycle and increasing each cycle until it reaches our chosen maximum value at the last cycle.

# Alternatively, C:\Users\Jonathan\OneDrive - Royal College of Surgeons in Ireland\COLOSSUS\Training Resources\Cost-Effectiveness and Decision Modeling using R Workshop _ DARTH\August_24\4_cSTM - time-dependent models_material shows you how to use a life-table, as does the material from the York course, but I really think there's no need to get that detailed in our own analysis.


## State rewards
# Costs and utilities  
c_H       <- 400   # cost of one cycle in healthy state
c_S       <- 1000  # cost of one cycle in sick state
c_D       <- 0     # cost of one cycle in dead state
c_trtA    <- 800   # cost of treatment A (per cycle) in healthy state
c_trtB    <- 1500  # cost of treatment B (per cycle) in healthy state
u_H       <- 1     # utility when healthy 
u_S       <- 0.5   # utility when sick
u_D       <- 0     # utility when dead
d_e       <- 0.03  # discount rate per cycle equal discount of costs and QALYs by 3%
d_c       <- 0.03  # discount rate per cycle equal discount of costs and QAL  

# Discount weight (equal discounting is assumed for costs and effects)
v_dwc <- 1 / (1 + d_c) ^ (0:n_cycles) 
v_dwe <- 1 / (1 + d_e) ^ (0:n_cycles) 
```

Discount rate for costs and utilities I've set as 3% here, I need to return to this and set it as 4%.
Also I set up 3 strategies, i.e. standard of care, treatment A and Treatment B, it would be very easy to change this to just 2 strategies above.
I need to change the health state names from "healthy, sick and dead" to "PFS, OS, Dead"





## Draw the state-transition cohort model

```{r}
m_P_diag <- matrix(0, nrow = n_states, ncol = n_states, dimnames = list(v_names_states, v_names_states))
m_P_diag["Healthy", "Sick" ]     = "" 
m_P_diag["Healthy", "Dead" ]     = ""
m_P_diag["Healthy", "Healthy" ]  = ""
m_P_diag["Sick"   , "Dead" ]     = ""
m_P_diag["Sick"   , "Sick" ]     = ""
m_P_diag["Dead"   , "Dead" ]     = ""
layout.fig <- c(2, 1)
plotmat(t(m_P_diag), t(layout.fig), self.cex = 0.5, curve = 0, arr.pos = 0.8,  
        latex = T, arr.type = "curved", relsize = 0.85, box.prop = 0.8, 
        cex = 0.8, box.cex = 0.7, lwd = 1)
```

# 04 Define and initialize matrices and vectors

## 04.1 Cohort trace

```{r}
## Initial state vector
# All starting healthy
v_s_init <- c("Healthy" = 1, "Sick" = 0, "Dead" = 0)  
v_s_init

## Initialize cohort trace for cSTM for all strategies
m_M_SoC <- matrix(0, 
                  nrow = (n_cycles + 1), ncol = n_states, 
                  dimnames = list(v_names_cycles, v_names_states))
# Store the initial state vector in the first row of the cohort trace
m_M_SoC[1, ] <- v_s_init
## Initialize cohort traces
m_M_trtA <- m_M_trtB <- m_M_SoC # structure and initial states remain the same
```

## 04.2 Transition probability matrix

```{r}
## Initialize transition probability matrix 
# all transitions to a non-death state are assumed to be conditional on survival 
m_P_SoC  <- matrix(0,
                   nrow = n_states, ncol = n_states,
                   dimnames = list(v_names_states, v_names_states)) # define row and column names
m_P_SoC
```

Fill in the transition probability matrix:

```{r}
## Standard of Care
# from Healthy
m_P_SoC["Healthy", "Healthy"] <- (1 - p_HD) * (1 - p_HS_SoC)
m_P_SoC["Healthy", "Sick"]    <- (1 - p_HD) *      p_HS_SoC
m_P_SoC["Healthy", "Dead"]    <-      p_HD

# from Sick
m_P_SoC["Sick", "Sick"] <- 1 - p_SD
m_P_SoC["Sick", "Dead"] <-     p_SD

# from Dead
m_P_SoC["Dead", "Dead"] <- 1

## Treatment A
m_P_trtA <- m_P_SoC
m_P_trtA["Healthy", "Healthy"] <- (1 - p_HD) * (1 - p_HS_trtA)
m_P_trtA["Healthy", "Sick"]    <- (1 - p_HD) *      p_HS_trtA

## Treatment B
m_P_trtB <- m_P_SoC
m_P_trtB["Healthy", "Healthy"] <- (1 - p_HD) * (1 - p_HS_trtB)
m_P_trtB["Healthy", "Sick"]    <- (1 - p_HD) *      p_HS_trtB
```

Check if transition probability matrices are valid.  

```{r}
# Check that transition probabilities are in [0, 1]
check_transition_probability(m_P_SoC,  verbose = TRUE)
check_transition_probability(m_P_trtA, verbose = TRUE)
check_transition_probability(m_P_trtB, verbose = TRUE)
# Check that all rows sum to 1
check_sum_of_transition_array(m_P_SoC,  n_states = n_states, verbose = TRUE)
check_sum_of_transition_array(m_P_trtA, n_states = n_states, verbose = TRUE)
check_sum_of_transition_array(m_P_trtB, n_states = n_states, verbose = TRUE)
```

# 05 Run Markov model

```{r}
for (t in 1:n_cycles){  # loop through the number of cycles
  m_M_SoC [t + 1, ] <- m_M_SoC [t, ] %*% m_P_SoC   # estimate the state vector for the next cycle (t + 1)
  m_M_trtA[t + 1, ] <- m_M_trtA[t, ] %*% m_P_trtA  # estimate the state vector for the next cycle (t + 1)
  m_M_trtB[t + 1, ] <- m_M_trtB[t, ] %*% m_P_trtB  # estimate the state vector for the next cycle (t + 1)
}
head(m_M_SoC)  # print the first few lines of the matrix
```

# 06 Compute and Plot Epidemiological Outcomes

## 06.1 Cohort trace

```{r}
matplot(m_M_SoC, type = 'l', 
        ylab = "Probability of state occupancy",
        xlab = "Cycle",
        main = "Cohort Trace", lwd = 3)  # create a plot of the data
legend("right", v_names_states, col = c("black", "red", "green"), 
       lty = 1:3, bty = "n")  # add a legend to the graph

# plot a vertical line that helps identifying at which cycle the prevalence of sick is highest
abline(v = which.max(m_M_SoC[, "Sick"]), col = "gray")  
```

## 06.2 Overall Survival (OS)

```{r}
v_os <- 1 - m_M_SoC[, "Dead"]    # calculate the overall survival (OS) probability
v_os <- rowSums(m_M_SoC[, 1:2])  # alternative way of calculating the OS probability   

plot(v_os, type = 'l', 
     ylim = c(0, 1),
     ylab = "Survival probability",
     xlab = "Cycle",
     main = "Overall Survival")  # create a simple plot showing the OS

# add grid 
grid(nx = n_cycles, ny = 10, col = "lightgray", lty = "dotted", lwd = par("lwd"), 
     equilogs = TRUE) 
```

## 06.2.1 Life Expectancy (LE)

```{r}
v_le <- sum(v_os)  # summing probability of OS over time  (i.e. life expectancy)
```

## 06.3 Disease prevalence

```{r}
v_prev <- m_M_SoC[, "Sick"]/v_os
plot(v_prev,
     ylim = c(0, 1),
     ylab = "Prevalence",
     xlab = "Cycle",
     main = "Disease prevalence")
```

# 07 Compute Cost-Effectiveness Outcomes

## 07.1 Mean Costs and QALYs

```{r}
# per cycle
# calculate expected costs by multiplying cohort trace with the cost vector for the different health states   
v_tc_SoC  <- m_M_SoC  %*% c(c_H, c_S, c_D)  
v_tc_trtA <- m_M_trtA %*% c(c_H + c_trtA, c_S, c_D)  
v_tc_trtB <- m_M_trtB %*% c(c_H + c_trtB, c_S, c_D)  

# calculate expected QALYs by multiplying cohort trace with the utilities for the different health states   
v_tu_SoC  <- m_M_SoC  %*% c(u_H, u_S, u_D)  
v_tu_trtA <- m_M_trtA %*% c(u_H, u_S, u_D) 
v_tu_trtB <- m_M_trtB %*% c(u_H, u_S, u_D) 
```

## 07.2 Discounted Mean Costs and QALYs

```{r}
# Discount costs by multiplying the cost vector with discount weights (v_dwc) 
tc_d_SoC  <-  t(v_tc_SoC)  %*% v_dwc
tc_d_trtA <-  t(v_tc_trtA) %*% v_dwc
tc_d_trtB <-  t(v_tc_trtB) %*% v_dwc

# Discount QALYS by multiplying the QALYs vector with discount weights (v_dwe)
tu_d_SoC  <-  t(v_tu_SoC)  %*% v_dwe
tu_d_trtA <-  t(v_tu_trtA) %*% v_dwe
tu_d_trtB <-  t(v_tu_trtB) %*% v_dwe

# Store them into a vector
v_tc_d <- c(tc_d_SoC, tc_d_trtA, tc_d_trtB)
v_tu_d <- c(tu_d_SoC, tu_d_trtA, tu_d_trtB)

# Dataframe with discounted costs and effectiveness
df_ce <- data.frame(Strategy = v_names_str,
                    Cost     = v_tc_d, 
                    Effect   = v_tu_d)
df_ce
```

## 07.3 Compute ICERs of the Markov model

```{r}
df_cea <- calculate_icers(cost       = df_ce$Cost,
                          effect     = df_ce$Effect,
                          strategies = df_ce$Strategy
                          )
df_cea
```

## 07.4 Plot frontier of the Markov model

```{r}
plot(df_cea, effect_units = "QALYs")
```








## References:




Useful Darth publications to cite when using this code:
 
- Jalal H, Pechlivanoglou P, Krijkamp E, Alarid-Escudero F, Enns E, Hunink MG. 
An Overview of R in Health Decision Sciences. Med Decis Making. 2017; 37(3): 735-746. 
https://journals.sagepub.com/doi/abs/10.1177/0272989X16686559

- Alarid-Escudero F, Krijkamp EM, Enns EA, Yang A, Hunink MGM Pechlivanoglou P,
Jalal H. Cohort State-Transition Models in R: A Tutorial. arXiv:200107824v2. 
2020:1-48. http://arxiv.org/abs/2001.07824

- Krijkamp EM, Alarid-Escudero F, Enns EA, Jalal HJ, Hunink MGM, Pechlivanoglou P. 
Microsimulation modeling for health decision sciences using R: A tutorial. 
Med Decis Making. 2018;38(3):400–22. 
https://journals.sagepub.com/doi/abs/10.1177/0272989X18754513
 
- Krijkamp EM, Alarid-Escudero F, Enns E, Pechlivanoglou P, Hunink MM, Jalal H. 
A Multidimensional Array Representation of State-Transition Model Dynamics. 
Med Decis Making. Online First https://doi.org/10.1177/0272989X19893973



